---
title: "Assignment 2_ Functions"
author: "Taylor Cook"
date: "2025-02-23"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Assignment 2

Think of a function that might be useful for environmental science and management data analysis. Keep it simple.

1. Code your function in R and store as a single .R file. Make sure you adequately comment and include some error checking.  

2. In Rmarkdown, generate some sample data and use this data to illustrate how your function might work. Then use this data to show what your function can do. You can be creative here - the illustration might be a graph or it could be a meaningful result (sum of yields). Just make sure you add enough text that we can understand what you did

3. Submit the .R file and the Rmarkdown file on Canvas

```{r}
# A function to calculate the volume of a tree, which can be used for forest management to calcualte the amount of wood that can be harvested from a tree or the economic value of a tree.
#inputs: diameter at breast height (dbh) and height of tree
#output: volume of tree
#volume = (pi/4) * dbh^2 * height

# pi/4 is the volume of a cylinder 
# dbh is the diameter of the tree at breast height
# height is the height of the tree.



#' Tree Volume Function
#'
#' @param dbh 
#' @param height 
#'
#' @return

tree_volume <- function(dbh, height) {
  # Check if any value in dbh is invalid
  if (any(dbh <= 0)) {
    stop("All DBH values must be greater than 0")
  }
  
  # Check if any value in height is invalid
  if (any(height <= 0)) {
    stop("All height values must be greater than 0")
  }
  
  # Compute tree volume (vectorized calculation)
  volume <- (pi / 4) * (dbh^2) * height
  return(volume)
}
```


```{r}
#generate some sample data
dbh <- c(10, 20, 30, 40, 50)
height <- c(50, 60, 70, 80, 90)

#use the function to calculate the volume of each tree

all_tree_volume <- tree_volume(dbh, height)
all_tree_volume
sum(all_tree_volume)

#for fun, we will calculate the economic value of the trees
#assume the price of wood is $100 per cubic meter
price_per_cubic_meter <- 100
economic_value <- all_tree_volume * price_per_cubic_meter
economic_value

sum(economic_value)


```


```{r}
#plot the volume of each tree

plot(dbh, all_tree_volume, xlab = "DBH", ylab = "Volume", main = "Volume of Trees")

```


```{r}
#plot the economic value of each tree

plot(dbh, economic_value, xlab = "DBH", ylab = "Economic Value", main = "Economic Value of Trees")

```


```{r}
# This is Jessica making a comment on your code :)
## Looks excellent! Thanks for helping me with Github :) 
```

